\documentclass[thesis.tex]{subfiles}
\begin{document}

\begin{abstract}

Descriptors are a tool used in computer vision for describing local structure in images. In this project we study the literature of descriptors, and develop our own descriptors based on this study. Our descriptors consist of localized kernel-based histograms, each describing parts of a region. Gradient orientation and shape index are used as histogram domains.

The descriptors are optimized and evaluated on two problems: image correspondence, where we test their ability to handle various image transformations, and pedestrian detection, where we test their ability to generalize the class of pedestrians. We compare the performance on each application against the SIFT and HOG descriptors. Extensive datasets of real world images are used.

Our descriptors perform marginally better than SIFT for image correspondence, and slightly worse than HOG for pedestrian detection. The addition of shape index to gradient orientation histograms is shown to be useful for pedestrian detection, but unnecessary for image correspondence.

\end{abstract}

\section*{Acknowledgements}
We would like to express our gratitude towards our supervisors Kim Steenstup Pedersen and Sune Darkner for their help and level of enthusiasm throughout this master's thesis. Furthermore we would like to thank the image group at DIKU for supplying us with access to the computational power needed in order to complete our work.

\section*{Collaboration details}
The work presented in this report is a joint effort, and hence both authors are responsible for the whole report.

\subbibliography

\end{document}
