\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath,amssymb,amsfonts}

\title{\bfseries{Study of higher order image descriptors}\\Article overview}
\author{
    Malte Stær Nissen \\\texttt{tgq958@alumni.ku.dk}
    \and
    Benjamin Michael Braithwaite \\ \texttt{cpg608@alumni.ku.dk}
    \\
    \\ \small{Supervisor: Kim Steenstrup Pedersen, \texttt{kimstp@diku.dk}}
    \\ \small{Co-supervisor: Sune Darkner, \texttt{darkner@diku.dk}}
    }

\begin{document}
\maketitle

This is an overview of the types of descriptors used in various articles.

\subsection{Crosier 2010 \cite{crosier2010using}}

Descriptor for 200x200 pixel texture regions. Consists of a histogram of 6 Basic Image Features (BIFs) built from 2-jet information at 4 different scales. ($6^4 = 1296$ bins)

\subsection{Dalal 2005 \cite{dalal2005histograms}}

Descriptor for 3x3 cell blocks of 6x6 pixel cells used for pedestrian detection. Each cell has a histogram of unsigned gradient orientations divided into 9 bins.

\subsection{Felzenszwalb 2008 \cite{felzenszwalb2008discriminatively}}

Descriptor for 2x2 cell blocks of 8x8 pixel cells used for general object detection. Each cell has a histogram of unsigned gradient orientations divided into 9 bins.

\subsection{Pedersen \cite{pedersenshape}}

Descriptor for 100x100 pixel regions of galaxies used to estimate specific star formation rate (sSFR). Consists of histograms of gradient orientation in 8 bins and shape index in 9 bins, at 8 different scale levels.

\bibliography{../bibliography}
\bibliographystyle{alpha}

\end{document}

