\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{hyperref}

\usepackage{amsmath,amssymb,amsfonts}

\title{\bfseries{Study of higher order image descriptors}\\Project description}
\author{
    Malte Stær Nissen \\\texttt{tgq958@alumni.ku.dk}
    \and
    Benjamin Michael Braithwaite \\ \texttt{cpg608@alumni.ku.dk}
    \\
    \\ \small{Supervisor: Kim Steenstrup Pedersen, \texttt{kimstp@diku.dk}}
    \\ \small{Co-supervisor: Sune Darkner, \texttt{darkner@diku.dk}}
    }

\begin{document}
\maketitle

\section{Problem statement}
Features/interest points are a central concept in computer vision. These
points mark areas of an image containing a significant amount of information
about the local structure. A visual descriptor is a representation of the
local structure of a feature. Such a descriptor can be modelled in numerous
ways depending on the application and desired properties. In this thesis we
will study state of the art visual descriptors, and develop and implement our
own descriptors based on this study. Furthermore we will evaluate our
descriptors against state of the art descriptors and discuss the results of the
evaluation.

\section{Motivation}
Computer scientists have for many years tried to make computers perceive
the world as humans do. This is done by performing analysis of real world
images. The analysis is often based on extraction of features in the images.
These features can then be be used for applications such as: the image
correspondence problem \cite{dahl2011finding}, object detection of pedestrians
\cite{felzenszwalb2008discriminatively}, content based image retrieval
\cite{smeulders2000content}, etc. The state of the art methods are however not
perfect and hence we will try to improve upon the accuracy of the methods.

\section{Proposed solution}
We will start by experimenting with a HOG descriptor extended to
various orders of differential information (like Jet descriptors à la
\cite{larsen2012jet}) and various parameters for the number of cells per
block, number of pixels per cell and number of channels per cell histogram.
Furthermore we will experiment with the combination of HOG and the shape
index \cite{koenderink1992surface} as well as different ways of processing
the higher order information. Since the higher order differential information
substantially increases the dimensionality we will experiment with PCA on the
information. We will also experiment with various ways of computing higher
order information and histograms, as well as how to compare histograms.

At first we will evaluate our
proposals on the image correspondence problem using the DTU robot image
dataset \cite{aanaes2012interesting}. Later we will select at least
one more of the following applications: Texture recognition, pedestrian
detection, general object detection using deformable parts models, medical
image registration, and content based image retrieval.

\section{Learning goals}
The following learning goals will be the success criteria for our
project.
\begin{itemize}
    \item Perform a literature study of visual descriptors and their
        applications.
    \item Describe the techniques used in state of the art descriptors such as
        HOG, SIFT, GLOH and DAISY.
    \item Develop, describe, and implement our own visual descriptor based on
        state of the art work within the field.
    \item Implement a library of competing state of the art descriptors.
    \item Implement a framework for evaluating descriptors on the image
        correspondence problem.
    \item Select and implement a framework for at least one (more if time
        permits) of the mentioned applications for evaluating descriptors.
    \item Perform a parameter study of our solutions.
    \item Perform an evaluation of our descriptor on the selected
        applications compared to state of the art descriptors.
    \item Discuss the results of the evaluation.
\end{itemize}

\section{Risk assessment}
The focus of the thesis is to gain insight into the field of descriptors
and develop our own non-trivial descriptor on the basis hereof. This makes
the risk of failure low since no requirements for the accuracy of the final
product are stated. There is however a risk of developing a solution which
performs worse than state of the art methods, but this will still have given
us significant insight into the field in question.

\section{Time schedule \small{(preliminary)}}

\begin{itemize}
    \item{\textbf{Literature study:}
        \textbf{Deadline:} Monday the 24th of February.}
    \item{\textbf{HOG and shape index:} Basic descriptor implementation and
        preliminary report section about descriptor techniques. \\
        \textbf{Deadline:} Wednesday the 5th of March.}
    \item{\textbf{Image correspondence implementation:} Image correspondence
        implementation and test with descriptor on DTU robot dataset. Chosen to
        get a simple initial estimate of the viability of our descriptor choices.
        Report section about the subject written. \\
        \textbf{Deadline:} Wednesday the 19th of March.}
    \item{\textbf{Unknown workload:} Further work will be determined
    continously.}
    \item{\textbf{Final experiments/results:}
        \textbf{Deadline:} Sunday the 3rd of August}
    \item{\textbf{Final report:}
        \textbf{Deadline:} Sunday the 17th of August}
\end{itemize}

\bibliography{../bibliography}
\bibliographystyle{alpha}

\end{document}

